---
title: "Customize your research website"
subtitle: "workflowr version `r packageVersion('workflowr')`"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Customize your research website}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

There are many ways to customize your research website. Below are some common
options.

## Adding project details

workflowr automatically creates many files when the project is first started. As
a first step for customizing your site, add the following information:

* Briefly describe your project in `analysis/index.Rmd`
* Share details about yourself in `analysis/about.Rmd`
* Copy the text for a license into `LICENSE` and state the name of the license in `analysis/license.Rmd`. See [A Quick Guide to Software Licensing for the Scientist-Programmer](http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002598) by Morin et al., 2012 for advice. If you're ambivalent, the MIT license is a standard choice
* Decide how authors should attribute your work by providing a citation in `CITATION`

## Changing the theme

The theme is defined in the file `analysis/_site.yml`. The default is cosmo, but
the rmarkdown package accepts multiple Bootstrap themes. These are listed in the
[rmarkdown documentation][rmd-themes]. Go to 
[bootswatch.com](http://bootswatch.com/) to compare the bootstrap themes. When
typing the theme, make sure it is all lowercase (e.g. spacelab, united, etc.).

If you are using RStudio (version > 1.0), you can quickly see how the theme
looks for one of your files. Open one of your R Markdown files, click Knit so
that it appears in the Viewer tab, and then edit `analysis/_site.yml`. Every time
you save an edit to the file, it will automatically re-render the R Markdown
file and update the result in the Viewer (thus you'll want to use a
quick-running file like `index.Rmd` when testing aesthetic changes).

Once you have chosen a theme, update the website by running the following:

```{r wflow-commit-theme}
wflow_commit(commit_files = "analysis/_site.yml", all = TRUE)
```

This commits `analysis/_site.yml`, re-builds every HTML file using the new
theme, and commits all the updated HTML pages.

[rmd-themes]: http://rmarkdown.rstudio.com/html_document_format.html#appearance_and_style

## Customize the navigation bar

The navigation bar appears on the top of each page. By default it includes links
to `index.html` (Home), `about.html` (About), `license.html` (License), and the 
workflowr repository (the GitHub icon). This is all specified in
`analysis/_site.yml`. To start, replace the URL to the workflowr GitHub
repository with the URL to your GitHub repository.

If you have other important pages, you can add them as well. For example, to add
the text "The main result" which links to `main-result.html`, you would add the
following:

```
    - text: "The main result"
      href: main-result.html
```

You can also create a drop-down menu from the navigation bar. See the [rmarkdown
documentation][navbar] for instructions.

Similar to changing the theme above, you will need to re-render each page of the
website (the navbar is embedded within each individual HTML file). Thus you
could run the same command as above:

```{r wflow-commit-navbar}
wflow_commit(commit_files = "analysis/_site.yml", all = TRUE)
```

[navbar]: http://rmarkdown.rstudio.com/rmarkdown_websites.html#site_navigation

## Add a links page

To make it easy to navigate to the pages of your website, you can manually add 
links either to `analysis/index.Rmd` or to the navigation bar defined in 
`analysis/_site.yml`. This can become tedious as the project grows to many 
files. To create a file called `analysis/results.Rmd` that contains links to 
each analysis file, run the workflowr function `create_links_page()`. By default
it sorts the files by their filename, but there are other options (see
`?create_links_page`).

Next add this page to the navigation bar in `analysis/_site.yml` to make it
easily accessible.

```
    - text: "Results"
      href: results.html
```

Lastly, since the navigation bar was changed, re-build the website using the following command:

```{r wflow-commit-links}
wflow_commit(commit_files = c("analysis/_site.yml", "analysis/results.Rmd"),
             all = TRUE)
```

Future updates of `results.Rmd` will not require re-building the entire website
since it is an independent page (as opposed to the navigation bar).

## Customize the footer

By default your workflowr website has the footer "This site was created with R
Markdown" added to the end of each page. This is specified in the file `analysis/include/footer.html` and this file is specified in `analysis/_site.yml`. To remove the footer entirely, delete the following lines:

```
    include:
      after_body: include/footer.html
```

To change the text that appears, directly edit the HTML file. Including
formatting (e.g. bold, hyperlink, etc.) requires knowing HTML.

This file is also how you can enable [disqus](https://disqus.com/) comments for
your website pages. To enable comments, follow these steps:

1. Start a disqus account
1. Add your disqus shortname to `footer.html`
1. Uncomment the disqus code in `footer.html`

To change the footer for every page of the website, run the following:

```{r wflow-commit-footer}
wflow_commit(commit_files = "analysis/include/footer.html", all = TRUE)
```

## Setup SSH keys

Using the https protocol to communicate with GitHub is tedious because it
requires entering your GitHub username and password. Using SSH keys for
authentication removes the password requirement. Follow these [GitHub
instructions][ssh] for creating SSH keys and linking them to your GitHub
account.

The easiest way to update your repository would be to delete your old version
(after pushing all the changes to GitHub of course) and then run `git clone` 
using the SSH URL. This will set the remote URL ("origin") to use SSH 
authentication. The other option would be to delete the HTTPS remote (`git
remote rm origin`) and then re-add "origin" using the SSH URL (`git remote add
origin <insert SSH URL>`).

[ssh]: https://help.github.com/articles/generating-an-ssh-key/
