---
title: "Getting started with workflowr"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```


The goal of the workflowr package is to make it easier for researchers to organize their projects and share their results with colleagues. If you are already writing R code to analyze data, and know the basics of Git and GitHub, you can start taking advantage of workflowr immediately. In a matter of minutes, you can create a research website like [this][demo01].

[demo01]: https://github.com/jdblischak/workflowr-demo01

## Setup

This tutorial assumes you have already R and Git installed, have a global
user.name and user.email set (you can run `git config -l` in the Terminal to
confirm), and have an account on GitHub. Rstudio is also convenient, especially
if you have a version greater than 1.0, but is not required. If you need
instructions to install these, [Software Carpentry][swc] maintains high-quality
installation instructions for use in its workshops. Follow the instructions for
your particular operating system to [install R][swc-r], [install Git][swc-git], and [configure Git][swc-git-config]. You can
obtain a GitHub account at https://github.com/.

workflowr is hosted in a [GitHub repository][repo]. Install it with devtools.

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("jdblischak/workflowr")
```

[swc]: https://software-carpentry.org
[swc-r]: https://swcarpentry.github.io/workshop-template/#r
[swc-git]: https://swcarpentry.github.io/workshop-template/#git
[swc-git-config]: http://swcarpentry.github.io/git-novice/02-setup/
[repo]: https://github.com/jdblischak/workflowr

As you use workflowr, if you find any unexpected behavior or think of an additional feature that would be nice to have, please open an Issue [here][issues]. When writing your bug report or feature request, please note the version of workflowr you are using (which you can obtain by running `packageVersion("workflowr")`).

[issues]: https://github.com/jdblischak/workflowr/issues

## Start the project

To start a new project, open R (or RStudio), load the workflowr package, and run the function `start_project`. This creates a directory called `new-project/` that contains all the files to get started, and also initializes a Git repository with the inital commit already made.

```{r start-project}
library("workflowr")
start_project("A new project", "new-project")
```

If you're using RStudio, you can choose `Open Project...` and select the file 
`new-project.Rproj`. This will set the correct working directory in the R 
console, switch the file navigator to the project, and configure the Git and 
Build panes. It will also reset the R environment, so you'll need to re-run 
`library("workflowr")`. If you're not using RStudio, set the working directory
to the root of the project directory.

The project directory contains multiple files and subdirectories. The goal is for these to help organize your project.

* Put raw data files in `data/`
* Put processed data files in `output/`
* Put long running scripts in `code/`. Ideally these steps can be automated using GNU Make or a similar tool
* Put R Markdown analysis files in `analysis/`. These should load data using relative paths, e.g. `d <- read.table("../output/clean-data.txt")`
* The rendered HTML files and other website code will automatically be put in `docs/` (described in the next section)

## Build the website

To build the website, run the function `render_site` in the R console, making sure that the argument `input` points to `analysis/`.

```{r render-site}
render_site(input = "analysis/")
```

This is much simpler if you are using RStudio version 1.0 or greater because you
do not have to worry about the current working directory. Click "Build Website"
in the Build pane (or use the keyboard shortcut Ctrl+Shift+B) to have RStudio
properly run `render_site` and load a preview of the website in the Viewer pane.

You now have a functioning website on your local computer. All the website files
are located in `docs/`. Open `docs/index.html` to view the website in your 
computer's web browser. Update the files `index.Rmd` and `about.Rmd` in the 
`analysis`, re-build the site, and refresh your browser to see your changes.

## Add a new analysis file

To start a new analysis, run the following command:

```{r open-rmd}
open_rmd("first-analysis.Rmd")
```

This performs multiple actions:

1. Creates a new file `analysis/first-analysis.Rmd` based on the workflowr R Markdown template (it doesn't overwrite the file if it already exists)
2. Sets the working directory to the `analysis/` directory
3. If you are using RStudio, opens the file for editing

Now you are ready to start writing! At the top of the file, edit the author,
title, and date. If you are using RStudio, press the Knit button to see a
preview in the Viewer pane. From the R console, you could run `render_site`
again, but this would re-render every R Markdown file. To only render those R
Markdown files that are new or have been updated, use the workflowr function
`make_site`.

```{r make-site}
make_site()
```

Check out your new file `docs/first-analysis.html`. Near the top you will see a line that says "Code version:" followed by an alphanumeric character string. This informs you which version of the code was used to create the file (this is explained more below in the section "Version the website").

In order to make it easier to navigate to your new file, you can include a link
to it on the main index page. For example, open `analysis/index.Rmd` and paste
the following line. This uses the Markdown syntax for creating a hyperlink (for
a quick reference guide in RStudio click "Help" -> "Markdown Quick Reference").
You specify the HTML version of the file since this is what is comprises the
website.

```
Click on this [link](first-analysis.html) to see my results.
```

Run `make_site` again to view the update.

## Version the website

One of the main goals of workflowr is to help make your research more
transparent and reproducible. An important component of this is using the
version control system Git. By including the unique identifier that Git assigns
a snapshot (or "commit" as Git calls it) at the top of each of your rendered
HTML files, you always know which version of the code produced the results.

To get started with workflowr, the main Git commands you will need to use are
`git add` and `git commit`. There are many ways to use Git: in the Terminal, in
the RStudio Git pane, or another Git graphical user interface (GUI) (see
[here](https://git-scm.com/download/gui/linux) for GUI options). This vignette will display the Git commands to run in the Terminal.

When you first started the project with `start_project`, workflowr automatically
made the first commit of all the project files. Since then you have updated some
of the R Markdown files, added a new file `first-analysis.Rmd`, and built the 
website in `docs/`. In order for the version of the code displayed at the top of
the file to be useful, it needs to refer to the exact state of the R Markdown
files when the HTML was rendered. To acheive this, you'll need to commit the R M
arkdown files and corresponding HTML in separate steps. First add and commit the
R Markdown files in `analysis/`.

```
git add analysis/*Rmd
git commit -m "Started my first analysis"
```

Second you'll need to re-render the corresponding HTML files so that they have
the latest commit identifier at the top of the file. However you can't use
`make_site` for this like you did during testing, because the R Markdown files
haven't changed. To re-build only those HTML files that correspond to the R
Markdown files in the previous commit, run the workflowr function `commit_site`.

```{r commit-site}
commit_site()
```

`commit_site` will render the R Markdown files and create a new commit with the
message "Build site.".

This is the general workflow.

1. Write code in the R Markdown files (For RStudio users: to quickly develop the code I recommend executing the code in the R console via Ctrl-Enter to send one line or Ctrl-Alt-C to execute the entire code chunk)
1. Run `make_site` to view the results as they will appear on the website
1. Go back to step 1 until you are satisfied with the result
1. Commit the R Markdown files in `analysis`
1. Run `commit_site` to build and commit the website

## Deploy the website
