---
title: "Getting started with workflowr"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The goal of the workflowr package is to make it easier for researchers to organize their projects and share their results with colleagues. If you are already writing R code to analyze data, and know the basics of Git and GitHub, you can start taking advantage of workflowr immediately. In a matter of minutes, you can create a research website like [this][demo01].

[demo01]: https://github.com/jdblischak/workflowr-demo01

## Setup

This tutorial assumes you have already R and Git installed, have a global
user.name and user.email set (you can run `git config -l` in the Terminal to
confirm), and have an account on GitHub. Rstudio is also convenient, especially
if you have a version greater than 1.0, but is not required. If you need
instructions to install these, [Software Carpentry][swc] maintains high-quality
installation instructions for use in its workshops. Follow the instructions for
your particular operating system to [install R][swc-r], [install Git][swc-git], and [configure Git][swc-git-config]. You can
obtain a GitHub account at https://github.com/.

workflowr is hosted in a [GitHub repository][repo]. Install it with devtools.

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("jdblischak/workflowr")
```

[swc]: https://software-carpentry.org
[swc-r]: https://swcarpentry.github.io/workshop-template/#r
[swc-git]: https://swcarpentry.github.io/workshop-template/#git
[swc-git-config]: http://swcarpentry.github.io/git-novice/02-setup/
[repo]: https://github.com/jdblischak/workflowr

As you use workflowr, if you find any unexpected behavior or think of an additional feature that would be nice to have, please open an Issue [here][issues]. When writing your bug report or feature request, please note the version of workflowr you are using.

```{r version}
packageVersion("workflowr")
```

[issues]: https://github.com/jdblischak/workflowr/issues

## Start the project

To start a new project, open R (or RStudio), load the workflowr package, and run the function `start_project`. This creates a directory called `new-project/` that contains all the files to get started, and also initializes a Git repository with the inital commit already made.

```{r start-project}
library("workflowr")
start_project("A new project", "new-project")
```

If you're using RStudio, you can choose `Open Project...` and select the file
`new-project.Rproj`. This will set the correct working directory in the R
console, switch the file navigator to the project, and configure the Git and
Build panes. It will also reset the R environment, so you'll need to re-run
`library("workflowr")`.

## Build the website

## Add a new analysis file

## Deploy the website

```{r teardown, include=FALSE}
unlink("new-project", recursive = TRUE)
```
