---
title: "Getting started with workflowr"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```


The goal of the workflowr package is to make it easier for researchers to organize their projects and share their results with colleagues. If you are already writing R code to analyze data, and know the basics of Git and GitHub, you can start taking advantage of workflowr immediately. In a matter of minutes, you can create a research website like [this][demo01].

[demo01]: https://jdblischak.github.io/workflowr-demo01/

## Setup

This tutorial assumes you have already R and Git installed, have a global
user.name and user.email set (you can run `git config -l` in the shell to
confirm), and have an account on GitHub. Rstudio is also convenient, especially
if you have a version greater than 1.0, but is not required. If you need
instructions to install these, [Software Carpentry][swc] maintains high-quality
installation instructions for use in its workshops. Follow the instructions for
your particular operating system to [install R][swc-r], [install Git][swc-git], and [configure Git][swc-git-config]. You can
obtain a GitHub account at https://github.com/.

workflowr is hosted in a [GitHub repository][repo]. Install it with devtools.

```{r install, eval=FALSE}
# install.packages("devtools")
devtools::install_github("jdblischak/workflowr", build_vignettes = TRUE)
```

[swc]: https://software-carpentry.org
[swc-r]: https://swcarpentry.github.io/workshop-template/#r
[swc-git]: https://swcarpentry.github.io/workshop-template/#git
[swc-git-config]: http://swcarpentry.github.io/git-novice/02-setup/
[repo]: https://github.com/jdblischak/workflowr

To see the available vignettes, run `browseVignettes("workflowr")`. This will
open a webpage with a link to each vignette. Currently the following vignettes are available (you are currently reading the first vignette):

Vignette title  | To open from R
------------- | -------------
Getting started with workflowr  | `vignette("getting-started", "workflowr")`
Customize your research website  | `vignette("customization", "workflowr")`
How the workflowr package works | `vignette("how-it-works", "workflowr")`

As you use workflowr, if you find any unexpected behavior or think of an additional feature that would be nice to have, please open an Issue [here][issues]. When writing your bug report or feature request, please note the version of workflowr you are using (which you can obtain by running `packageVersion("workflowr")`).

[issues]: https://github.com/jdblischak/workflowr/issues

## Start the project

To start a new project, open R (or RStudio), load the workflowr package, and run the function `start_project`. This creates a directory called `new-project/` that contains all the files to get started, and also initializes a Git repository with the inital commit already made.

```{r start-project}
library("workflowr")
start_project("A new project", "new-project")
```

If you're using RStudio, you can choose `Open Project...` and select the file 
`new-project.Rproj`. This will set the correct working directory in the R 
console, switch the file navigator to the project, and configure the Git and 
Build panes. It will also reset the R environment, so you'll need to re-run 
`library("workflowr")`. If you're not using RStudio, set the working directory
to the root of the project directory.

The project directory contains multiple files and subdirectories. The goal is for these to help organize your project.

* Put raw data files in `data/`
* Put processed data files in `output/`
* Put long running scripts in `code/`. Ideally these steps can be automated using GNU Make or a similar tool
* Put R Markdown analysis files in `analysis/`. These should load data using relative paths, e.g. `d <- read.table("../output/clean-data.txt")`
* The rendered HTML files and other website code will automatically be put in `docs/` (described in the next section)

## Build the website

To build the website, run the function `render_site` in the R console, making sure that the argument `input` points to `analysis/`.

```{r render-site}
render_site(input = "analysis/")
```

This is much simpler if you are using RStudio version 1.0 or greater because you
do not have to worry about the current working directory. Click "Build Website"
in the Build pane (or use the keyboard shortcut Ctrl+Shift+B) to have RStudio
properly run `render_site` and load a preview of the website in the Viewer pane.

You now have a functioning website on your local computer. All the website files
are located in `docs/`. Update the files `index.Rmd` and `about.Rmd` in 
`analysis/` and re-build the site to see your changes. If you aren't using
RStudio, open `docs/index.html` to view the website in your computer's web
browser and refresh your browser after running `render_site` to see your
changes.

## Add a new analysis file

To start a new analysis, run the following command:

```{r open-rmd}
open_rmd("first-analysis.Rmd")
```

This performs multiple actions:

1. Creates a new file `analysis/first-analysis.Rmd` based on the workflowr R Markdown template (it doesn't overwrite the file if it already exists)
2. Sets the working directory to the `analysis/` directory
3. If you are using RStudio, opens the file for editing

Now you are ready to start writing! At the top of the file, edit the author, 
title, and date. If you are using RStudio, press the Knit button to render the
file and see a preview in the Viewer pane. From the R console, you could run
`render_site` again (equivalent to clicking "Build Website" in RStudio), but
this would re-render every R Markdown file. To only render those R Markdown
files that are new or have been updated, use the workflowr function `make_site`.

```{r make-site}
make_site()
```

Check out your new file `docs/first-analysis.html`. Near the top you will see a
line that says "Code version:" followed by an alphanumeric character string.
This informs you which version of the code was used to create the file (this is
explained more below in the section [Version the
website](#version-the-website)).

In order to make it easier to navigate to your new file, you can include a link
to it on the main index page. For example, open `analysis/index.Rmd` and paste
the following line. This uses the Markdown syntax for creating a hyperlink (for
a quick reference guide in RStudio click "Help" -> "Markdown Quick Reference").
You specify the HTML version of the file since this is what comprises the
website.

```
Click on this [link](first-analysis.html) to see my results.
```

Click Knit (or run `make_site` again) to render the file.

## Version the website

One of the main goals of workflowr is to help make your research more
transparent and reproducible. An important component of this is using the
version control system Git. By including the unique identifier that Git assigns
a snapshot (or "commit" as Git calls it) at the top of each of your rendered
HTML files, you always know which version of the code produced the results.

To get started with workflowr, the main Git commands you will need to use are 
`git add` and `git commit`. There are many ways to use Git: in the shell, in 
the RStudio Git pane, or another Git graphical user interface (GUI) (see 
[here](https://git-scm.com/download/gui/linux) for GUI options). This vignette 
will display the Git commands to run in the shell (you can quickly open the
Terminal to the correct working directoy from RStudio by choosing "Tools" ->
"Shell...").

When you first started the project with `start_project`, workflowr automatically
made the first commit of all the project files. Since then you have updated some
of the R Markdown files, added a new file `first-analysis.Rmd`, and built the 
website in `docs/`. In order for the version of the code displayed at the top of
the file to be useful, it needs to refer to the exact state of the R Markdown 
files when the HTML was rendered. To acheive this, you'll need to commit the R 
Markdown files and corresponding HTML in separate steps. First add and commit 
the R Markdown files in `analysis/`.

```
# Run in the shell
git add analysis/*Rmd
git commit -m "Started my first analysis"
```

Second you'll need to re-render the corresponding HTML files so that they have
the latest commit identifier at the top of the file. However you can't use
`make_site` for this like you did during testing, because the R Markdown files
haven't changed. To re-build only those HTML files that correspond to the R
Markdown files in the previous commit, run the workflowr function `commit_site`.

```{r commit-site}
commit_site()
```

`commit_site` will render the R Markdown files and create a new commit with the
message "Build site.".

This is the general workflow.

1. Write code in the R Markdown files (For RStudio users: to quickly develop the code I recommend executing the code in the R console via Ctrl-Enter to send one line or Ctrl-Alt-C to execute the entire code chunk)
1. Run `make_site` to view the results as they will appear on the website
1. Go back to step 1 until you are satisfied with the result
1. Commit the R Markdown files in `analysis`
1. Run `commit_site` to build and commit the website

You only need to run `render_site`, which re-renders every single R Markdown
file, if you are updating the aesthetics of your website (e.g. anytime you make
edits to `_analysis/_site.yml`).

## Deploy the website

At this point you have built a version-controlled website that exists on your
local computer. The next step is to put your code on GitHub so that it can serve
your website online. To do this, login to your account on GitHub and create a
new repository following these [instructions][new-repo]. Make sure you do not
add an automatically-generated README, .gitignore, or license (these are
important, but workflowr already creates them for you). Next you will copy the
Git commands under the heading "…or push an existing repository from the command
line". Using the hypothetical GitHub username "myname" and hypothetical 
repository name "myproject"^[The name of the repository on GitHub does not need
to be identical to the directory name of your local Git repo; however, it is
convenient to have them match since this is the default behavior of `git clone`
when copying your repo to a another computer], the two lines will look like the
following:

```
# Run in the shell
git remote add origin https://github.com/myname/myproject.git
git push -u origin master
```

The first line creates the alias "origin" that points to your remote repository
on GitHub^["origin" is the convential name, but could be anything you wanted]. 
The second pushes your current commit history to GitHub. You will be prompted to
enter your GitHub username and password for authentication after you run `git
push`.

Now that your code is on GitHub, you need to tell GitHub that you want the files
in `docs/` to be published as a website. Go to Settings -> GitHub Pages and 
choose "master branch docs/ folder" as the Source
([instructions][publish-docs]). Using the hypothetical names above, the
repository would be hosted at the URL `https://myname.github.io/myproject/`^[It 
may take a few minutes for the site to be rendered.]. If you scroll back down to
the GitHub Pages section of the Settings page, you can click on the URL there.

Note that even if you create a private GitHub repository, the website it creates
is still public. If you're not ready to publish your results online, you can 
always wait and activate GitHub Pages later. In the meantime, you'll still have 
a version-controlled, organized set of results for your project. However, the
risk that someone that doesn't have the link to your website is able to find it
is very low. Search engines prioritize the results by how many other sites link
to a site, so your website will not be high in the results even if you search 
for very specific terms. Thus if you only share the URL to your results with
your close collaborators, and request that they not share it widely, your
website is effectively private. That being said, being truly scooped in science 
is rare (at best) and openly sharing your work will help establish your 
expertise in the field (and furthermore establishes your priority), so you
should consider keeping both your code and website public.

[new-repo]: https://help.github.com/articles/creating-a-new-repository/
[publish-docs]: https://help.github.com/articles/configuring-a-publishing-source-for-github-pages/#publishing-your-github-pages-site-from-a-docs-folder-on-your-master-branch

## Further reading

* For advice on using R Markdown files to organize your analysis, read the
chapter [R Markdown workflow](http://r4ds.had.co.nz/r-markdown-workflow.html) in
the book [R for Data Science](http://r4ds.had.co.nz/) by Garrett Grolemund and
Hadley Wickham
