---
title: "Migrating an existing project to use workflowr"
subtitle: "workflowr version `r packageVersion('workflowr')`"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Migrating an existing project to use workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

## Introduction

This vignette is for those users that already have an existing project and wish 
to incorporate workflowr to create a research website. Migrating an existing 
project to use workflowr varies from straightforward to difficult depending on 
the scenario and your comfort level with Git. This vignette assumes that you
have the background knowledge of workflowr explained in the Getting Started
vignette. Even if you have no need for a new workflowr project, please run
through that vignette first as an exercise to familiarize yourself with the
workflowr philosophy and functions.

```{r getting-started}
vignette("getting-started", "workflowr")
```

## Scenario: I have a collection of R Markdown files (no version control)

## Scenario: I have a collection of R Markdown files (with version control)

## Scenario: I have a collection of R Markdown files with lots of other supporting files (no version control)

## Scenario: I have a collection of R Markdown files with lots of other supporting files (with version control)

## Scenario: I have an R package (no version control)

## Scenario: I have an R package (with version control)

## Scenario: I have an ashlar project
