---
title: "Migrating an existing project to use workflowr"
subtitle: "workflowr version `r packageVersion('workflowr')`"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Migrating an existing project to use workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

## Introduction

This vignette is for those users that already have an existing project and wish 
to incorporate workflowr to create a research website. Migrating an existing 
project to use workflowr varies from straightforward to difficult depending on 
the scenario and your comfort level with Git. This vignette assumes that you
have the background knowledge of workflowr explained in the Getting Started
vignette. Even if you have no need for a new workflowr project, please run
through that vignette first as an exercise to familiarize yourself with the
workflowr philosophy and functions.

```{r getting-started}
vignette("getting-started", "workflowr")
```

## Scenario: I have a collection of R Markdown files

If you have a collection of R Markdown files, but no version control or other 
files, the most straightfoward solution would be to create a workflowr project 
in a new directory and then move the R Markdown files to the `analysis/` 
subdirectory. In the hypothetical example below, the original R Markdown files 
are located in the directory `~/projects/misc/` and the workflowr project will
be created in the new directory `~/projects/new-project/`.

```{r}
library("workflowr")
wflow_start("~/projects/new-project")
file.copy(from = glob2rx("~/projects/misc/*Rmd"),
          to = "~/projects/new-project/analysis")
```

## Scenario: I have a collection of R Markdown files and other project infrastructure

If your project already has lots of infrastructure, it is most convenient to add
the workflowr files directory to your already existing directory. This is
controlled with the argument `existing`. In the hypothetical example below, the
existing project is located at `~/projects/mature-project/`.

```{r}
library("workflowr")
wflow_start("~/projects/mature-project", existing = TRUE)
```

The above command will add the workflowr files to your existing project and also
commit them to version control (it will initialize a Git repo if it doesn't 
already exist). If you'd prefer to not use version control for your project or
you'd prefer to commit the workflowr files yourself manually, you can set `git =
FALSE` (this is also useful if you want to first test to see what would happen
without committing the results).

To add your R Markdown files to the research website, you can move them to the 
subdirectory `analysis/` (note you can do this before or after running 
`wflow_start`).

Lastly, by default `wflow_start` will not overwrite your existing files (e.g. if
you already have a `README.md`). If you'd prefer to overwrite your files with
the default workflowr files, set `overwrite = TRUE`.

## Scenario: I have an R package

If your project is organized as an R package, you can still add a website using
workflowr. In the hypothetical example below, the
existing package is located at `~/projects/my-package/`.

```{r}
library("workflowr")
wflow_start("~/projects/my-package", existing = TRUE)
```

The above command will add the workflowr files to your existing project and also
commit them to version control (it will initialize a Git repo if it doesn't 
already exist). If you'd prefer to not use version control for your project or
you'd prefer to commit the workflowr files yourself manually, you can set `git =
FALSE` (this is also useful if you want to first test to see what would happen
without committing the results).

You'll want R to ignore the workflowr directories when building the R pacakge.
Thus add the following to the `.Rbuildignore` file:

```
^analysis$
^docs$
^data$
^code$
^output$
```

If your primary purpose for creating a website to accompany your package is to 
share the package documentation, please check out the package [pkgdown][]. It 
creates a website from the vignettes and function documentation files (i.e. the 
Rd files in `man/`). In contrast, if the purpose of the website is to
demonstrate results you obtained using the package, use workflowr.

[pkgdown]: https://github.com/hadley/pkgdown

## Scenario: I have an ashlar project

First, make sure you are in the "master" branch. If you have any HTML files in
`analysis/` remove them using Git at the command line.

```bash
git rm analysis/*html
```

Additionally, remove `chunk-options.R`.

```bash
git rm analysis/chunk-options.R
```

Next add the workflowr files. In the hypothetical example below, the existing
package is located at `~/projects/ashlar-project/`.

```{r}
library("workflowr")
wflow_start("~/projects/ashlar-project", existing = TRUE)
```

Lastly, you'll need to update each of your R Markdown files. Make the following substitutions:

```{r verbatim-inline-r-code-hack, include=FALSE}
# I had to use a hack to get the inline R code to show verbatim as opposed to
# being evaluated.
#
# http://stackoverflow.com/a/20419016
```

ashlar:

```
**Code version:** `r '\x60r system(\"git log -1 --format=\'%H\'\", intern = TRUE)\x60'`
```

workflowr:

```
**Code version:** `r '\x60r workflowr::extract_commit(".", 1)$sha1\x60'
```

ashlar:

<pre><code>```{r chunk-options, include=FALSE}
source("chunk-options.R")
```</code></pre>

workflowr:

<pre><code>```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```</code></pre>

<pre><code>```{r knitr-opts-chunk, include=FALSE}
```</code></pre>


ashlar:

<pre><code>```{r info}
sessionInfo()
```</code></pre>

workflowr:

<pre><code>```{r session-info}
```</code></pre>
