---
title: "Hosting workflowr websites using GitLab"
subtitle: "workflowr version `r utils::packageVersion('workflowr')`"
author: "Luke Zappia, John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Frequently asked questions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## What is in this vignette?

By default workflowr assumes that the project will be hosted on GitHub, but this
is not always the case. Users may prefer to use another service or have a
private git repository hosting instance. This vignette details some of the steps
required to host a workflowr project on GitLab. Similar steps will be required
for other platforms but some of the specifics will be different.

As these steps are not currently integrated into workflowr this vignette assumes
a basic knowledge of using git from a command line interface.

## Step 0: Set up a project

The first thing we need to do is set up the project we want to host. We can do
this by following the first few steps of the instructions in the
"Getting started" vignette. When you get to the section [Deploy the website](wflow-01-getting-started.html#deploy-the-website), follow
the rest of the steps in this vignette.

## Step 1: Set up a remote repository

Log in to the GitLab instance you want to use and create a repository to host
your project. You will need to know your user name and the repository name for
the following steps (here we are going to use "myname" and "myproject") as well
as a URL for the hosting instance. The example below assumes you are using
GitLab.com. If instead you are using a custom instance of GitLab, you will need
to change the value for the argument `domain` accordingly ^[For example, the
University of Chicago hosts a GitLab instance for its researchers at
https://git.rcc.uchicago.edu/, which would require setting `domain =
"git.rcc.uchicago.edu"`].

```{r wflow-git-remote, eval=FALSE}
wflow_git_remote(remote = "origin", user = "myname", repo = "myproject",
                 domain = "gitlab.com")
```

We now want to connect our local git repository to the one we have just created
on GitLab:

```{r wflow-git-push, eval=FALSE}
wflow_git_push()
```

If this step has worked correctly you should be able to refresh your GitLab
page and see all the files in your workflowr project.

## Step 2: Change the build directory

By default workflowr stores produced HTML files in the `docs` directory but
GitLab can only host files that are in a directory called `public`. To fix this
we need to tell workflowr to change the directory name by editing the
`_site.yml` file in the `analysis` directory. We want to change the line that
says `output_dir: "../docs"` to `output_dir: "../public"`. The command below
uses `sed` to do this but you could also do it in RStudio or your favourite text
editor.

```
sed -i'.bk' s/docs/public/ analysis/_site.yml
```

We already have a directory called `docs` so we also need to tell git that we
want to change the name:

```
git mv docs/ public/
```

## Step 3: Add a GitLab CI config file

While we can easily set up web page hosting for GitHub using just the website
interface for GitLab we need to do a bit more work. GitLab using their
continuous integration system to manage building and hosting website so
we need to provide a config file with instructions on how to do this. This
file should be called `.gitlab-ci.yml` and be placed at the root of our project
directory. Because our HTML pages have already been built by workflowr we don't
actually need GitLab to run any code for us so our config file can be quite
simple. Here is an example of what the config file should look like:

```
pages:
    stage: deploy
    script:
        - echo 'Nothing to do...'
    artifacts:
        paths:
            - public
    only:
        - master
```

> **NOTE:** If you are using a private GitLab instance you might also need to
set up a GitLab runner and assign it to your project. Instructions on how to
do that are outside the scope of this vignette.

At this stage we should have everything we need to host our project. Let's
commit what we have and push it to GitLab.

```
git add .gitlab-ci.yml
git add analysis/_site.yml
git commit -m "Switch to hosting with GitLab"
git push origin master
```

We should only have to do this manually the first time, after that the normal
`wflow_publish` function should work.

If this has worked correctly should now be able to see your GitLab hosted
workflowr project website! To find the link to the website go to your GitLab
repository page and select "Pages" under the "Settings" menu on the left. If
the link is missing or doesn't work the job to build the website may not have
run successfully. You can check this by clicking on the "CI / CD" menu.

## Step 4: Fixing the links

If you look carefully at your new website you should notice that many of the
links that are usually there are missing. This is a result of us not using the
`wflow_git_remote` command earlier. To fix these links we need to add the
following line to the `_workflowr.yml` file.

```
github: GITLAB_REPOSITORY_URL
```

Here GITLAB_REPOSITORY_URL should be a URL to your project GitLab repository,
not the URL for your project website.

To fix the links in our exisiting web pages we need to rebuild them. We can
force workflowr to do this by setting the `republish` option in the 
`wflow_publish` function to `TRUE`.

```
wflow_publish("_workflowr.yml", republish = TRUE)
```

Check your website to make sure you have links to previous versions of versions
of code, figures etc. Note that the links to previous HTML documents on GitLab
will not work with the current version of workflowr.
