---
title: "Hosting workflowr websites using GitLab"
subtitle: "workflowr version `r utils::packageVersion('workflowr')`"
author: "Luke Zappia, John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Hosting workflowr websites using GitLab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## What is in this vignette?

By default workflowr assumes that the project will be hosted on GitHub, but this
is not always the case. Users may prefer to use another service or have a
private git repository hosting instance. This vignette details how to host a
workflowr project on GitLab. For more details, see the documentation for [GitLab
Pages](https://docs.gitlab.com/ee/ci/yaml/README.html#pages). Similar steps will
be required for other platforms but some of the specifics will be different.

## Step 0: Set up a project

The first thing we need to do is set up the project we want to host. We can do
this by following the first few steps of the instructions in the
"Getting started" vignette. When you get to the section [Deploy the website](wflow-01-getting-started.html#deploy-the-website), follow
the rest of the steps in this vignette.

## Step 1: Create a remote repository on GitLab

Log in to the GitLab instance you want to use and create a repository to host
your project (follow this [link](https://gitlab.com/projects/new) if you are
using GitLab.com). We recommend setting the project to be Public so that others
can inspect the code behind your results and extend your work.

## Step 2: Configure your local workflowr project to use GitLab

You will need to know your user name and the repository name for the following
steps (here we are going to use "myname" and "myproject") as well as a URL for
the hosting instance. The example below assumes you are using GitLab.com. If
instead you are using a custom instance of GitLab, you will need to change the
value for the argument `domain` accordingly ^[For example, the University of
Chicago hosts a GitLab instance for its researchers at
https://git.rcc.uchicago.edu/, which would require setting `domain =
"git.rcc.uchicago.edu"`].

```{r wflow-use-gitlab, eval=FALSE}
wflow_use_gitlab(username = "myname", repository = "myproject")
```

The function `wflow_use_gitlab()` automates all the local configuration
necessary to use GitLab. It changes the website directory from `docs/` to
`public/`, it creates the GitLab-specific configuration file `.gitlab-ci.yml`
with the necessary settings, and it connects the local Git repository to
communicate with the remote respository on GitLab.

## Step 3: Republish the analyses

In order for the correct URLs to past versions to be inserted into the HTML
pages, republish the analyses with `wflow_publish()`.

```
wflow_publish(republish = TRUE)
```

## Step 4: Push to GitLab

As a final step, push the workflowr project to GitLab (you will be prompted for
your GitLab username and password):

```{r wflow-git-push, eval=FALSE}
wflow_git_push()
```

If this step has worked correctly you should be able to refresh your GitLab page
and see all the files in your workflowr project. You can view your site at
https://myname.gitlab.io/myproject/, replacing with your username and project
(note it may take a minute for the site to be deployed).
