---
title: "Sharing common code across analyses"
subtitle: "workflowr version `r utils::packageVersion('workflowr')`"
author: "Tim Trice, John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Sharing common code across analyses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

During the course of a project, you may want to repeat a similar analysis
across multiple R Markdown files. To avoid duplicated code across your files
(which is difficult to update), there are multiple strategies you can use to
share common code:

1. To share R code like function definitions, you can put this code in an R
script and import it in each file with the function `source()`

1. To share common R Markdown text and code chunks, you can use [child documents](https://yihui.name/knitr/demo/child/)

1. To share common templates, you can use the function `knitr::knit_expand()`

Each of these strategies is detailed below, with a special emphasis on how to
use them within the workflowr framework. In order to source scripts or use child
documents, it is suggested you use the [here][] package, which helps to locate
the root directory of your project regardless of the directory your script or
analysis file is, making sourcing documents cleaner.

[here]: https://cran.r-project.org/package=here

## Overview of directories

First, a quick overview of the directories in a workflowr project. This is
critical for importing these shared files.

In a standard R Markdown file, the code is executed in the directory where the R
Markdown file is saved. Thus any paths to files in the R Markdown file should be
relative to this directory. However, the directory where the code is executed,
referred to as the "knit directory" in the workflowr documentation, can be
configured. The default for a new workflowr project is to run the code in the
root of the workflowr project (this is defined in the file `_workflowr.yml`; see
`?wflow_html` for configuration details). Thus any filepaths should be relative
to the root of the project. As an example, if you have shared R functions
defined in the file `~/Desktop/myproject/code/common.R`, the relative filepath
from the root of the project directory would be `"code/common.R"`.

## Share R code with source()

If you have R code you want to re-use across multiple R Markdown files, the most
straightforward option is to save this code in an R script, e.g.
`code/functions.R`.

Then in each R Markdown file that needs to use the code defined in that file,
you can use `source()` to load it. If the code in your workflowr project is
executed in the root of the project directory (which is the default behavior for
new workflowr projects), then you would add the following chunk:

````
`r ''````{r shared-code}
source("code/functions.R")
```
````

On the other hand, if you have changed the value of `knit_root_dir` in the file
`_workflowr.yml`, you need to ensure that the filepath to the R script is
relative to this directory. For example, if you set `knit_root_dir: "analysis"`,
you would use this code chunk:

````
`r ''````{r shared-code}
source("../code/functions.R")
```
````

To avoid having to figure out the correct relative path (or having to update it
in the future if you were to change `knit_root_dir`), you can use `here::here()`
as it is always based off the project root. Additionally, it will help
readability when using child documents as discussed below.

````
`r ''````{r shared-code}
source(here::here("code/functions.R"))
```
````

## Share child documents with chunk option

To share text and code chunks across R Markdown files, you can use [child
documents](https://yihui.name/knitr/demo/child/), a feature of the [knitr][]
package.

[knitr]: https://cran.r-project.org/package=knitr

Here is a example of a simple R Markdown file that you can use to test this
feature. Note that it contains an H2 header, some regular text, and a code
chunk.

````
## Header in child document

Text in child document.

`r ''````{r child-code-chunk}
str(mtcars)
```
````

You can save this child document anywhere in the workflowr project with one
critical exception: it cannot be saved in the R Markdown directory (`analysis/`
by default) with the file extension `.Rmd` or `.rmd`. This is because workflowr
expects every R Markdown file in this directory to be a standalone analysis that
has a 1:1 correspondence with an HTML file in the website directory (`docs/` by
default). We recommend saving child documents in a subdirectory of the R
Markdown directory, e.g. `analysis/child/ex-child.Rmd`.

To include the content of the child document, you can reference it using
`here::here()` in your chunk options.

````
`r ''````{r parent, child = here::here("analysis/child/ex-child.Rmd")}
```
````

However, this fails if you wish to include plots in the code chunks of the child
documents. It will not generate an error, but the plot will be missing ^[The
reason for this is very technical and requires more understanding of how
workflowr is implemented than is necessary to use it effectively in the majority
of cases. Whenever workflowr builds an R Markdown file, it first copies it to a
temporary directory so that it can inject extra code chunks that implement some
of its reproducibility features. The figures in the child documents end up being
saved there and then lost.]. In a situation like this, you would want to
generate the plot within the parent R Markdown file or use
`knitr::knit_expand()` as described in the next section.

## Share templates with knit_expand()

If you need to pass parameters to the code in your child document, then you can
use `knitr::knit_expand()`. Also, this strategy has the added benefit that it
can handle plots in the child document. However, this requires setting
`knit_root_dir: "analysis"` in the file `_workflowr.yml` for plots to work
properly.

Below is an example child document with one variable to be expanded: `{{title}}`
refers to a species in the iris data set. The value assigned will be used to
filter the iris data set and label the section, chunk, and plot. We will refer
to this file as `analysis/child/iris.Rmd`.

````

## {{title}}

`r ''````{r plot_{{title}}}
iris %>%
  filter(Species == "{{title}}") %>%
  ggplot() +
  aes(x = Sepal.Length, y = Sepal.Width) +
  geom_point() +
  labs(title = "{{title}}")
```
````

To generate a plot using the species `"setosa"`, you can expand the child
document in a hidden code chunk:

````
`r ''````{r, include = FALSE}
src <- knitr::knit_expand(file = here::here("analysis/child/iris.Rmd"),
                          title = "setosa")
```
````

and then later knit it using an inline code expression^[Before calling
`knitr::knit()`, you'll need to load the dplyr and ggplot2 packages to run the
code in this example child document.]:

`` `r
knitr::knit(text = unlist(src))` ``

The convenience of using `knitr::knit_expand()` gives you the flexibility to
generate multiple plots along with custom headers, figure labels, and more. For
example, if you want to generate a scatter plot for each Species in the `iris`
datasets, you can call `knitr::knit_expand()` within a `lapply()` or
`purrr::map()` call:

````
`r ''````{r, include = FALSE}
src <- lapply(
  sort(unique(iris$Species)), 
  FUN = function(x) {
    knitr::knit_expand(
      file = here::here("analysis/child/iris.Rmd"), 
      title = x
    )
  }
)
```
````

This example code loops through each unique `iris$Species` and sends it to the
template as the variable `title`. `title` is inserted into the header, the chunk
label, the `dplyr::filter()`, and the title of the plot. This generates three
plots with custom plot titles and labels while keeping your analysis flow clean
and simple.

Remember to insert `knitr::knit(text = unlist(src))` in an inline R expression
as noted above to knit the code in the desired location of your main document.

Read the `knitr::knit_expand()` vignette for more information.

```{r knit-expand-vignette}
vignette("knit_expand", package = "knitr")
```
