---
title: "Sharing common code across analyses"
subtitle: "workflowr version `r utils::packageVersion('workflowr')`"
author: "Tim Trice, John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Customize your research website}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

During the course of a project, you may want to repeat a similar analysis
across multiple R Markdown files. To avoid duplicated code across your files
(which is difficult to update), there are multiple strategies you can use to
share common code:

1. To share R code like function definitions, you can put this code in an R
script and import in each file with the function `source()`

1. To share common R Markdown text and code chunks, you can use child documents

1. To share common templates, you can use the knitr function `knit_expand()`

Each of these strategies is detailed below, with a special emphasis on how to
use them within the workflowr framework.

## Overview of directories

First, a quick overview of the directories in a workflowr project. This is
critical for importing these shared files.

In a standard R Markdown file, the code is executed in the directory where the R
Markdown file is saved. Thus any paths to files in the R Markdown file should be
relative to this directory. However, the directory where the code is executed,
referred to as the "knit directory" in the workflowr documentation, can be
configured. The default for a new workflowr project is to run the code in the
root of the workflowr project (this is defined in the file `_workflowr.yml`; see
`?wflow_html` for configuration details). Thus any filepaths should be relative
to the root of the project. As an example, if you have shared R functions
defined in the file `~/Desktop/myproject/code/common.R`, the relative filepath
from the root of the project directory would be `"code/common.R"`.

## Share R code with source()

## Share child documents with chunk option

https://yihui.name/knitr/demo/child/

## Share templates with knit_expand()

```{r knit-expand-vignette}
vignette("knit_expand", package = "knitr")
```

