---
title: "Migrating an existing project to use workflowr"
subtitle: "workflowr version `r utils::packageVersion('workflowr')`"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Migrating an existing project to use workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

## Introduction

This vignette is for those users that already have an existing project and wish 
to incorporate workflowr to create a research website. Migrating an existing 
project to use workflowr varies from straightforward to difficult depending on 
the scenario and your comfort level with Git. This vignette assumes that you
have the background knowledge of workflowr explained in the [Getting started][vig-start]
vignette. Even if you have no need for a new workflowr project, please run
through that vignette first as an exercise to familiarize yourself with the
workflowr philosophy and functions.

```{r getting-started}
vignette("wflow-01-getting-started", "workflowr")
```

[vig-start]: wflow-01-getting-started.html

## Scenario: I have a collection of R Markdown files

If you have a collection of R Markdown files, but no version control or other 
files, the quickest solution is to use the function `wflow_quickstart()`. The
code below 1) starts a new workflowr project in `~/projects/new-project/`,
2) copies the existing Rmd files in `~/projects/misc/` to the `analysis/`
subdirectory of the new project, 3) builds and commits the website, and 4)
configures the project to use GitHub (which is why the GitHub username is
required).

```{r}
library("workflowr")
wflow_quickstart("~/projects/misc/*Rmd", username = "<github-username>",
                 directory = "~/projects/new-project/")
```

Alternatively, you can manually perform each step to migrate your existing
analysis by starting a workflowr project in a new directory and then moving the
R Markdown files to the `analysis/` subdirectory. In the hypothetical example
below, the original R Markdown files are located in the directory
`~/projects/misc/` and the workflowr project will be created in the new
directory `~/projects/new-project/`.

```{r}
library("workflowr")
# Create project directory and change working directory to this location
wflow_start("~/projects/new-project")
# Copy the files to the analysis subdirectory of the workflowr project
file.copy(from = Sys.glob("~/projects/misc/*Rmd"), to = "analysis")
```

Next run `wflow_build()` to see if your files run without error. Lastly, build
and commit the website using `wflow_publish()`:

```{r}
wflow_publish("analysis/*Rmd", "Publish analysis files")
```

When you are ready to share the results online, you can run `wflow_use_github()`
or `wflow_use_gitlab()`.

## Scenario: I have a collection of R Markdown files and other project infrastructure

If your project already has lots of infrastructure, it is most convenient to add
the workflowr files directory to your already existing directory. This is
controlled with the argument `existing`. In the hypothetical example below, the
existing project is located at `~/projects/mature-project/`.

```{r}
library("workflowr")
wflow_start("~/projects/mature-project", existing = TRUE)
```

The above command will add the workflowr files to your existing project and also
commit them to version control (it will initialize a Git repo if it doesn't 
already exist). If you'd prefer to not use version control for your project or
you'd prefer to commit the workflowr files yourself manually, you can set `git =
FALSE` (this is also useful if you want to first test to see what would happen
without committing the results).

By default `wflow_start()` will not overwrite your existing files (e.g. if
you already have a `README.md`). If you'd prefer to overwrite your files with
the default workflowr files, set `overwrite = TRUE`.

To add your R Markdown files to the research website, you can move them to the 
subdirectory `analysis/` (note you can do this before or after running 
`wflow_start()`).

Next run `wflow_build()` to see if your files run without error. Lastly, build
and commit the website using `wflow_publish()`:

```{r}
wflow_publish("analysis/*Rmd", "Publish analysis files")
```

## Scenario: I have an R package

If your project is organized as an R package, you can still add a website using
workflowr. In the hypothetical example below, the
existing package is located at `~/projects/my-package/`.

```{r}
library("workflowr")
wflow_start("~/projects/my-package", existing = TRUE)
```

The above command will add the workflowr files to your existing project and also
commit them to version control (it will initialize a Git repo if it doesn't 
already exist). If you'd prefer to not use version control for your project or
you'd prefer to commit the workflowr files yourself manually, you can set `git =
FALSE` (this is also useful if you want to first test to see what would happen
without committing the results).

You'll want R to ignore the workflowr directories when building the R package.
Thus add the following to the `.Rbuildignore` file:

```
^analysis$
^docs$
^data$
^code$
^output$
^_workflowr.yml$
```

Furthermore, to prevent R from compressing the files in `data/` (which is
harmless but time-consuming), you can set `LazyData: false` in the file
`DESCRIPTION`. However, if you do want to distribute data files with your R
package, you'll need to instead rename the workflowr subdirectory and update the
R Markdown files to search for files in the updated directory name (and also
update `.Rbuildignore` to ignore this new directory and not `data/`). Then you
can save the data files to distribute with the package in `data/`. For more
details, see the relevant sections in the CRAN manual [Writing R
Extensions][data-in-packages] and Hadley's [R Packages][r-pkgs-data].

[data-in-packages]: https://cran.r-project.org/doc/manuals/r-release/R-exts.html#Data-in-packages
[r-pkgs-data]: http://r-pkgs.had.co.nz/data.html

If your primary purpose for creating a website to accompany your package is to 
share the package documentation, please check out the package [pkgdown][]. It 
creates a website from the vignettes and function documentation files (i.e. the 
Rd files in `man/`). In contrast, if the purpose of the website is to
demonstrate results you obtained using the package, use workflowr.

[pkgdown]: https://github.com/r-lib/pkgdown

## Scenario: I have an old worklowr project

If you created a workflowr project prior to its release to CRAN (any version
less than 1.0), you can migrate the project to use the 1.0+ features by running
`wflow_update()`:

```{r}
# Install the latest version from CRAN
install.packages("workflowr")
library("workflowr")
# Preview the files that would be changed
wflow_update()
# Update the files
wflow_update(dry_run = FALSE)
# Republish your analyses with the latest features
wflow_publish(republish = TRUE)
```

A few notes:

1. You do *not* have to upgrade if you don't want. To continue using the Beta
version of workflowr, follow the [workflowrBeta installation
instructions][beta-install].

1. For your updated project, the code in the R Markdown files will continue to be
executed from the R Markdown directory (typically `analysis/`). However, if you
create a new workflowr project, the default directory for executing the code
will be the root directory of the project (see `?wflow_html` for background and
instructions for changing this behavior).

[beta-install]: https://github.com/jdblischak/workflowrBeta#quick-start
