---
title: "Migrating an existing project to use workflowr"
subtitle: "workflowr version `r packageVersion('workflowr')`"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Migrating an existing project to use workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

## Introduction

This vignette is for those users that already have an existing project and wish 
to incorporate workflowr to create a research website. Migrating an existing 
project to use workflowr varies from straightforward to difficult depending on 
the scenario and your comfort level with Git. This vignette assumes that you
have the background knowledge of workflowr explained in the [Getting started][vig-start]
vignette. Even if you have no need for a new workflowr project, please run
through that vignette first as an exercise to familiarize yourself with the
workflowr philosophy and functions.

```{r getting-started}
vignette("wflow-01-getting-started", "workflowr")
```

[vig-start]: wflow-01-getting-started.html

## Scenario: I have a collection of R Markdown files

If you have a collection of R Markdown files, but no version control or other 
files, the most straightforward solution would be to create a workflowr project 
in a new directory and then move the R Markdown files to the `analysis/` 
subdirectory. In the hypothetical example below, the original R Markdown files 
are located in the directory `~/projects/misc/` and the workflowr project will
be created in the new directory `~/projects/new-project/`.

```{r}
library("workflowr")
# Create project directory and change working directory to this location
wflow_start("~/projects/new-project")
# Copy the files to the analysis subdirectory of the workflowr project
file.copy(from = Sys.glob("~/projects/misc/*Rmd"), to = "analysis")
```

Next update your R Markdown files to use the workflowr template (you can
first run `wflow_update()` with no arguments to preview the potential changes).

```{r}
wflow_update(dry_run = FALSE)
```

Lastly, build and commit the website using `wflow_publish()`:

```{r}
wflow_publish("analysis/*Rmd", "Publish analysis files")
```

## Scenario: I have a collection of R Markdown files and other project infrastructure

If your project already has lots of infrastructure, it is most convenient to add
the workflowr files directory to your already existing directory. This is
controlled with the argument `existing`. In the hypothetical example below, the
existing project is located at `~/projects/mature-project/`.

```{r}
library("workflowr")
wflow_start("~/projects/mature-project", existing = TRUE)
```

The above command will add the workflowr files to your existing project and also
commit them to version control (it will initialize a Git repo if it doesn't 
already exist). If you'd prefer to not use version control for your project or
you'd prefer to commit the workflowr files yourself manually, you can set `git =
FALSE` (this is also useful if you want to first test to see what would happen
without committing the results).

By default `wflow_start()` will not overwrite your existing files (e.g. if
you already have a `README.md`). If you'd prefer to overwrite your files with
the default workflowr files, set `overwrite = TRUE`.

To add your R Markdown files to the research website, you can move them to the 
subdirectory `analysis/` (note you can do this before or after running 
`wflow_start()`).

Next run `wflow_update()` to convert the R Markdown files to use the workflowr
template:

```{r}
wflow_update(dry_run = FALSE)
```

Lastly, build and commit the website using `wflow_publish()`:

```{r}
wflow_publish("analysis/*Rmd", "Publish analysis files")
```

## Scenario: I have an R package

If your project is organized as an R package, you can still add a website using
workflowr. In the hypothetical example below, the
existing package is located at `~/projects/my-package/`.

```{r}
library("workflowr")
wflow_start("~/projects/my-package", existing = TRUE)
```

The above command will add the workflowr files to your existing project and also
commit them to version control (it will initialize a Git repo if it doesn't 
already exist). If you'd prefer to not use version control for your project or
you'd prefer to commit the workflowr files yourself manually, you can set `git =
FALSE` (this is also useful if you want to first test to see what would happen
without committing the results).

You'll want R to ignore the workflowr directories when building the R package.
Thus add the following to the `.Rbuildignore` file:

```
^analysis$
^docs$
^data$
^code$
^output$
```

If your primary purpose for creating a website to accompany your package is to 
share the package documentation, please check out the package [pkgdown][]. It 
creates a website from the vignettes and function documentation files (i.e. the 
Rd files in `man/`). In contrast, if the purpose of the website is to
demonstrate results you obtained using the package, use workflowr.

[pkgdown]: https://github.com/hadley/pkgdown

## Scenario: I have an ashlar project

If your project uses the [ashlar][] template, you'll need to follow multiple
steps to convert it to a valid workflowr setup.

[ashlar]: https://github.com/jhsiao999/ashlar

First, make sure you are in the "master" branch. Also, you should run the
following Git commands from the base of the project directory.

```bash
git checkout master
```

If you have any HTML files in `analysis/` remove them using Git at the command
line.

```bash
git rm analysis/*html
```

Next remove the other ashlar infrastructure files.

```bash
git rm analysis/chunk-options.R analysis/Makefile analysis/_output.yaml
git rm -r analysis/libs analysis/include analysis/figure
```

Modify `.gitignore` so that it ignores HTML and figure files in `analysis/` only
by changing `*html` to `analysis/*html` and `*png` to `analysis/figure`.
Afterwards add the file.

```bash
nano .gitignore
git add .gitignore
```

And move the RStudio project file from `analysis/` to the base of the project.

```bash
git mv analysis/*Rproj .
```

Commit these changes.

```bash
git commit -m "Removed ashlar infrastructure."
```

Next open R to add the workflowr files. In the hypothetical example below, the
existing package is located at `~/projects/ashlar-project/`.

```{r}
library("workflowr")
wflow_start("~/projects/ashlar-project", existing = TRUE)
```

This will generate a few warning messages stating that the directory already
contains a `.git` directory and a `.gitignore` file. These can be safely ignored
because ashlar projects are expected to be using Git for version control.

Lastly, you'll need to update your R Markdown files to use the workflowr
template, which you can do using the function `wflow_update()`:

```{r convert-ashlar}
wflow_update()
# Run with `dry_run = FALSE` to apply the updates
# wflow_update(dry_run = FALSE)
```

This will update every R Markdown file, with the exception of `index.Rmd`, 
`about.Rmd`, and `license.Rmd`. It will also fail to automatically convert any
ashlar files that do not follow the template (e.g. they do not have the
`sessionInfo()` chunk), so you will have to manually update these files. Also,
workflowr no longer includes the date and SHA1 in the organizational files like
`index.Rmd`, so you can remove this manually if you like.

Now you should have a functioning workflowr project! From here you should be
able to make a few more small edits and then follow the workflowr steps to
create your website:

* Edit the navbar in `analysis/_site.yml` so that the GitHub URL points to your repository (and commit the change)
* Edit the README (and potentially other places, including on GitHub) so that the URL to the website no longer contains the trailing "analysis" in the name (and commit the change(s))
* Run `wflow_publish("analysis/*Rmd", "Publish files using workflowr"))` to build and commit the website files
* Push the changes to GitHub
* Change the Settings so that the website is served from "master branch /docs folder"
* Optionally remove the gh-pages branch ([GitHub instructions][gh-remove-branch])

[gh-remove-branch]: https://github.com/blog/1377-create-and-delete-branches
