---
title: "Getting started with workflowr"
subtitle: "workflowr version `r packageVersion('workflowr')`"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Getting started with workflowr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, cache=FALSE, include=FALSE}
library("knitr")
.tmp <- tempfile("wflow-01-getting-started-")
.project <- file.path(.tmp, "myproject")
dir.create(.project, recursive = TRUE)
opts_knit$set(root.dir = .project)
```

The goal of the workflowr package is to make it easier for researchers
to organize their projects and share their results with colleagues. If
you are already writing R code to analyze data, and know the basics of
Git and GitHub, you can start taking advantage of workflowr
immediately. In a matter of minutes, you can create a research website
like [this][demo01].

[demo01]: https://jdblischak.github.io/workflowr-demo01/

This tutorial assumes you have already followed the "Quick start"
instructions in the
[README](https://jdblischak.github.io/workflowr/index.html). Specifically,
you should have already installed R, Git, and workflowr; set the Git
variables user.name and user.email (you can run `git config -l` in the
shell to confirm); and created an account on GitHub.

## Overview

A workflowr project has two key components:

1. An R Markdown-based website. This consists of a configuration file
(`_site.yml`), a collection of R Markdown files, and their
corresponding HTML files.

2. A Git repository. Git is a [version control system][vcs] that
helps track code development^[There are many ways to use Git: in the
shell, in the RStudio Git pane, or another Git graphical user
interface (GUI) (see [here](https://git-scm.com/download/gui/linux)
for GUI options). When necessary, this vignette will display the Git
commands to run in the shell (you can quickly open the Terminal to the
correct working directory from RStudio by choosing "Tools" ->
"Shell...").]. All the Git files are in a hidden directory `.git` at
the root of the project.

One of the main goals of workflowr is to help make your research more
transparent and reproducible. This is achieved by including the unique
identifier that Git assigns a snapshot of your code (or "commit" as
Git calls it) at the top of each of your HTML files, so you always
know which version of the code produced the results.

[vcs]: https://en.wikipedia.org/wiki/Version_control

## Start the project

To start a new project, open R (or RStudio), load the workflowr
package, and run the function `wflow_start()`. The code below creates
a directory called `myproject/` that contains all the files to get
started. It also changes the working directory to `myproject/` and
initializes a Git repository with the initial commit already made.

```{r start-project, echo=3:4}
setwd(.tmp)
unlink(.project, recursive = TRUE)
library("workflowr")
wflow_start("myproject")
```

> **pc-comment-5:** One important step that is missing here: where is
> directory `myproject` being created? The concept of the "working
> directory" in R is missing from many novice R programmers, and I
> guarantee that many people will be wondering: where was my project
> created? Make sure that the user first sets the working directory
> appropriately, either using `setwd` or using the menu options in
> RStudio. You should also suggest running `getwd` to make sure that
> we are in the correct directory before running `wflow_start`.

> **pc-comment-16:** You didn't mention that this command
> automatically changes the working directory; another good
> opportunity to check this with `getwd`.

If you're using RStudio, you can choose `Open Project...` and select
the file `myproject.Rproj`. This will set the correct working
directory in the R console, switch the file navigator to the project,
and configure the Git and Build panes. It will also reset the R
environment, so you'll need to re-run `library("workflowr")`.

> **pc-comment-6:** This part about RStudio is confusing. Is this
> instead of running `wflow_start`, or after running it? Does it do
> anything useful? What happens if you don't do this? I would suggest
> giving this info after first explaining the result of running
> `wflow_start`.

`wflow_start()` creates the following directory structure in
`myproject/`:

```
myproject/
├── analysis/
│   ├── about.Rmd
│   ├── chunks.R
│   ├── include/
│   │   └── footer.html
│   ├── index.Rmd
│   ├── license.Rmd
│   ├── README.md
│   └── _site.yml
├── CITATION
├── code/
│   ├── README.md
│   ├── script.py*
│   ├── script.R*
│   └── script.sh*
├── data/
│   └── README.md
├── docs/
├── LICENSE
├── myproject.Rproj
├── output/
│   └── README.md
└── README.md
```

> **pc-comment-8:** At this point I suggest saything something like
> this: "At this point, you have a minimal but complete workflowr
> project; that is, you have all the files needed to use the main
> workflowr commands andpublish a research website (admittedly, this
> research website is not very interesting!). Later on, as you get
> more comfortable with the basic setup, you can modify and add to
> the initial file structure."
>
> You might also want to comment on the overall rationale for this
> structure; that is, it is intended to help organize the files that
> will be commonly included in a data analysis project.

The two **required** subdirectories are `analysis/` and `docs/`. These
directories should never be removed from the workflowr project.

> **pc-comment-7:** What about removing the files within these
> directories? Is it okay to remove them? 

* `analysis/`: This directory contains all the source R Markdown files
for implementing the data analyses for your project. It also contains
a special R Markdown file, `index.Rmd`, that does not contain any R
code, but will be used to generate `index.html`, the homepage for your
website. In addition, this directory contains important configurations
files like `_site.yml` and `chunks.R`.

> **pc-comment-10:** If `index.Rmd` is important, I suppose we
> shouldn't delete it?

* `docs/`: This directory contains all the HTML files for your
website. Some of the HTML files are built from the R Markdown files in
`analysis/`. Furthermore, any figures created by the R Markdown files
are saved here. Each of these figures is saved according to the
following pattern: `docs/figure/<insert Rmd filename>/<insert chunk
name>-#.png`, where `#` corresponds to which of the plots the chunk
generated (since one chunk can produce an arbitrary number of plots).

> **pc-comment-12:** Does it *only* contain HTML files built from R
> Markdown files? See my change above.

The **optional** directories are `data/`, `code/`, and `output/`.
These directories are suggestions for organizing your data analysis
project, but can be removed if you do not find them useful.

* `data/`: This directory is for raw data files.

> **pc-comment-11:** Why only "raw"? 

* `code/`: This directory is for code and scripts that might not be
appropriate as R Markdown notebooks (e.g. for pre-processing the data, or
for long-running code).

* `output/`: This directory is for processed data files and other
outputs generated from the code and data. For example, scripts in
`code` that pre-process raw data files from `data/` should save the
processed data files in `output/`.

> **pc-comment-9:** You didn't say anything about the other files
> (e.g., `myproject.Rproj`, `CITATION`). Which are essential, and
> which can be deleted?

## Build the website

You will notice that the `docs/` directory is currently empty. That is
because we have not yet generated the website from the `analysis/`
files. This is what we will do next.

> **pc-comment-13**: This is a good opportunity to make sure that the
> user is in the correct working directory.

To build the website, run the function `wflow_build()` in the R
console:

```{r wflow-build}
wflow_build()
```

This commands builds all the R Markdown files in `analysis/` and saves
the corresponding HTML files in `docs/`. It sets the same seed before
running every file so that any function that generates random data
(e.g. permutations) is reproducible. Furthermore, each file is built
in its own external R session to avoid any potential conflicts between
analyses (e.g. accidentally sharing a variable with the same name
across files).

> **pc-comment-14:** Do you "build an R Markdown file" or "build an
> HTML file from an R Markdown file"? What is the correct usage here?

The default action of `wflow_build()` is to behave similar to a
[Makefile](https://swcarpentry.github.io/make-novice) (`make = TRUE`
is the default), i.e. it only builds R Markdown files that have been
modified more recently than their corresponding HTML files. Thus if
you run it again, no files are built.

> **pc-comment-15:** I think it is useful to mention Makefiles here,
> but it would be good to give an explanation for people that are
> unfamiliar with the idea; e.g., "`wflow_build()` only does the work
> that is necessary; it will automatically compare the modification
> dates of the analysis and HTML files, and only re-build an HTML file
> when the corresponding analysis file has changed.

```{r wflow-build-no-action}
wflow_build()
```

To view the site, run `wflow_view()`, which by default opens the file
`docs/index.html` in your browser:

```{r wflow-view, eval=FALSE}
wflow_view()
```

This is how you can preview any changes you make right on your local machine. Go
ahead and edit the files `index.Rmd`, `about.Rmd`, and `license.Rmd` to describe
your project. Run `wflow_build()` to re-build the HTML files and `wflow_view()`
to preview the changes (alternatively refresh the page in your browser).

```{r edit-files, include=FALSE}
for (f in file.path("analysis", c("index.Rmd", "about.Rmd", "license.Rmd"))) {
  cat("\nedit\n", file = f, append = TRUE)
}
```

## Publish the website

workflowr makes an important distinction between R Markdown files that are
published versus unpublished. A published file is included in the website
online; whereas, the HTML file of an unpublished R Markdown file is only able to
be viewed on the local computer. Since the project was just started, there are
no published files. To view the status of the workflowr project, run
`wflow_status()`.

```{r wflow-status}
wflow_status()
```

This alerts us that our project has 3 R Markdown files, and they are all
unpublished ("Unp"). Furthermore, it instructs how to publish them: use
`wflow_publish()`. The first argument to `wflow_publish()` is a character vector
of the R Markdown files to publish. The second is a message that will recorded
by the version control system Git when it commits (i.e. saves a snapshot of)
these files. The more informative the commit message the better (so that future
you knows what you were trying to accomplish).

```{r wflow-publish}
wflow_publish(c("analysis/index.Rmd", "analysis/about.Rmd", "analysis/license.Rmd"),
              "Publish the initial files for myproject")
```

`wflow_publish()` reports the 3 steps it took:

* **Step 1:** Commits the 3 R Markdown files using the custom commit message

* **Step 2:** Builds the HTML files using `wflow_build()`

* **Step 3:** Commits the 3 HTML files plus the files that specify the style of
the website (e.g. CSS and Javascript files)

Performing these 3 steps ensures that the HTML files are always in sync with the
latest versions of the R Markdown files. Performing these steps manually would
be tedious and error-prone (e.g. an HTML file may have been built with an
outdated version of an R Markdown file). However, `wflow_publish()` makes it
trivial to keep the pages of your site in sync.

Now when you run `wflow_status()`, it reports that all the files are published
and up-to-date.

```{r wflow-status-post-publish}
wflow_status()
```

## Deploy the website

At this point you have built a version-controlled website that exists on your
local computer. The next step is to put your code on GitHub so that it can serve
your website online. To do this, login to your account on GitHub and create a
new repository following these [instructions][new-repo]. Make sure you do not
add an automatically-generated README, .gitignore, or license (these are
important, but workflowr already creates them for you). Next you will copy the
Git commands under the heading "…or push an existing repository from the command
line". Using the hypothetical GitHub username "myname" and hypothetical
repository name "myproject"^[The name of the repository on GitHub does not need
to be identical to the directory name of your local Git repo; however, it is
convenient to have them match since this is the default behavior of `git clone`
when copying your repo to a another computer], the two lines will look like the
following:

```
# Run in the shell
git remote add origin https://github.com/myname/myproject.git
git push -u origin master
```

The first line creates the alias "origin" that points to your remote repository
on GitHub^["origin" is the conventional name, but could be anything you wanted].
The second pushes your current commit history to GitHub. You will be prompted to
enter your GitHub username and password for authentication after you run `git
push`. To view or update your remote repositories, use the function
`wflow_remotes()`.

Now that your code is on GitHub, you need to tell GitHub that you want the files
in `docs/` to be published as a website. Go to Settings -> GitHub Pages and
choose "master branch docs/ folder" as the Source
([instructions][publish-docs]). Using the hypothetical names above, the
repository would be hosted at the URL `https://myname.github.io/myproject/`^[It
may take a few minutes for the site to be rendered.]. If you scroll back down to
the GitHub Pages section of the Settings page, you can click on the URL there.

[new-repo]: https://help.github.com/articles/creating-a-new-repository/
[publish-docs]: https://help.github.com/articles/configuring-a-publishing-source-for-github-pages/#publishing-your-github-pages-site-from-a-docs-folder-on-your-master-branch

## Add a new analysis file

Now that you have a functioning website, the next step is to start analyzing
data! To start a new analysis called `first-analysis.Rmd`, use `wflow_open()`:

```{r create-file, include=FALSE}
# Because devtools_shims overrides system.file, wflow_open can't work when
# building documentation with devtools::document. Thus create a blank file so
# that it doesn't try to copy the template via rmarkdown::draft.
file.create("analysis/first-analysis.Rmd")
```

```{r open-rmd, echo=1}
wflow_open("first-analysis.Rmd")
setwd("..")
```

This performs multiple actions:

1. Creates a new file `analysis/first-analysis.Rmd` based on the workflowr R Markdown template (it doesn't overwrite the file if it already exists)
2. Sets the working directory to the `analysis/` directory
3. If you are using RStudio, opens the file for editing

Now you are ready to start writing! At the top of the file, edit the author,
title, and date. Where it says "Add your analysis here", add some code chunks to
experiment. If you are using RStudio, press the Knit button to build the file
and see a preview in the Viewer pane. Alternatively from the R console, you can
run `wflow_build()` again and `wflow_view("first-analysis.Rmd")` (these
functions can be run from the base directory of your project or any
subdirectory).

Check out your new file `first-analysis.html`. Near the top you will see a line
that says "Code version:" followed by an alphanumeric character string. This
informs you which version of the code was used to create the file. It also
automatically inserts the date when the HTML was built.

In order to make it easier to navigate to your new file, you can include a link
to it on the main index page. First open `analysis/index.Rmd` (you can use your
filesystem navigator or `wflow_open("index.Rmd")`). Second paste
the following line into `index.Rmd`:

```
Click on this [link](first-analysis.html) to see my results.
```

```{r edit-index, include=FALSE}
cat("\nClick on this [link](first-analysis.html) to see my results.\n",
    file = "analysis/index.Rmd", append = TRUE)
```

This uses the Markdown syntax for creating a hyperlink (for a quick reference
guide in RStudio click "Help" -> "Markdown Quick Reference"). You specify the
HTML version of the file since this is what comprises the website. Click Knit
(or run `wflow_build()` again) to check that the link works.

Now run `wflow_status()` again. As expected, two files need attention.
`index.Rmd` has status "Mod" for modified. This means it is a published file
that has subsequently been modified. `first-analysis.Rmd` has status "Scr" for
Scratch. This means not only it the HTML not published, but the R Markdown file
is not yet being tracked by Git.

```{r wflow-status-newfile, echo=2}
setwd("analysis")
wflow_status()
setwd("..")
```

To publish the new analysis and the updated index page, again use
`wflow_publish()`:

```{r wflow-publish-newfile, echo=2:3}
setwd("analysis")
# Assuming working directory is `analysis/`. Run getwd() to confirm
wflow_publish(c("index.Rmd", "first-analysis.Rmd"),
              "Add my first analysis")
setwd("..")
```

Lastly, push the changes to GitHub to deploy these latest changes to the website.

```
# Run in the shell
git push origin master
```

## The workflow

This is the general workflow:

1. Open a new or existing file with `wflow_open()`

1. Perform your analysis in the R Markdown file (For RStudio users: to quickly
develop the code I recommend executing the code in the R console via Ctrl-Enter
to send one line or Ctrl-Alt-C to execute the entire code chunk)

1. Run `wflow_build()` and `wflow_view()` to view the results as they will
appear on the website (alternatively press the Knit button in RStudio)

1. Go back to step 2 until you are satisfied with the result

1. Run `wflow_publish()` to commit the source files (R Markdown files or other
files in `code/`, `data/`, and `output/`), build the HTML files, and commit the
HTML files

1. Push the changes to GitHub with `git push origin master`

This ensures that the "Code version:" inserted into each HTML file corresponds
to the state of the Git repository at the time the HTML was built.

The only exception to this workflow is if you are updating the aesthetics of
your website (e.g. anytime you make edits to `analysis/_site.yml`). In this case
you'll want to update all the published HTML files, regardless of whether or not
their corresponding R Markdown files have been updated. To republish every HTML
page, run `wflow_publish()` with `republish = TRUE`. This behavior is only
previewed below by specifying `dry_run = TRUE`.

```{r republish, echo=2:3}
setwd("analysis")
# Assuming working directory is `analysis/`. Run getwd() to confirm
wflow_publish("_site.yml", republish = TRUE, dry_run = TRUE)
setwd("..")
```

## Next steps

To learn more about workflowr, you can read the following vignettes:

* [Customize your research website](wflow-02-customization.html)
* [Migrating an existing project to use workflowr](wflow-03-migrating.html)
* [How the workflowr package works](wflow-04-how-it-works.html)
* [Frequently asked questions](wflow-05-faq.html)

## Further reading

* For advice on using R Markdown files to organize your analysis, read the
chapter [R Markdown workflow](http://r4ds.had.co.nz/r-markdown-workflow.html) in
the book [R for Data Science](http://r4ds.had.co.nz/) by Garrett Grolemund and
Hadley Wickham
