---
title: "Customize your research website"
subtitle: "workflowr version `r utils::packageVersion('workflowr')`"
author: "John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Customize your research website}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r chunk-options, include=FALSE}
library("knitr")
opts_chunk$set(eval = FALSE)
```

There are many ways to customize your research website. Below are some common
options.

## Adding project details

workflowr automatically creates many files when the project is first started. As
a first step for customizing your site, add the following information:

* Briefly describe your project in `analysis/index.Rmd`
* Share details about yourself in `analysis/about.Rmd`
* State a software license in `analysis/license.Rmd`. See [A Quick Guide to
Software Licensing for the Scientist-Programmer][morin2012] by Morin et al.,
2012 for advice. If you're ambivalent, the MIT license is a standard choice.

[morin2012]: http://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1002598

## Changing the theme

The theme is defined in the file `analysis/_site.yml`. The default is cosmo, but
the rmarkdown package accepts multiple Bootstrap themes. These are listed in the
[rmarkdown documentation][rmd-themes]. Go to 
[bootswatch.com](http://bootswatch.com/) to compare the bootstrap themes. When
typing the theme, make sure it is all lowercase (e.g. spacelab, united, etc.).

When experimenting with different themes, you'll want to build a fast-running
file, e.g. likely `analysis/index.Rmd`, instead of rebuilding the entire site
every time. Click the RStudio Knit button or run `wflow_build()` in the R
console to preview each theme:

```
wflow_build("analysis/index.Rmd")
```

Once you have chosen a theme, update the website by running the following:

```{r wflow-publish-theme}
wflow_publish("analysis/_site.yml", "Change the theme", republish = TRUE)
```

This commits `analysis/_site.yml`, re-builds every previously published HTML
file using the new theme, and commits all the republished HTML pages.

[rmd-themes]: http://rmarkdown.rstudio.com/html_document_format.html#appearance_and_style

## Customize the navigation bar

The navigation bar appears on the top of each page. By default it includes links
to `index.html` (Home), `about.html` (About), `license.html` (License), and the 
workflowr repository (the GitHub icon). This is all specified in
`analysis/_site.yml`. To start, replace the URL to the workflowr GitHub
repository with the URL to your GitHub repository.

If you have other important pages, you can add them as well. For example, to add
the text "The main result" which links to `main-result.html`, you would add the
following:

```
    - text: "The main result"
      href: main-result.html
```

You can also create a drop-down menu from the navigation bar. See the [rmarkdown
documentation][navbar] for instructions.

Similar to changing the theme above, you will need to re-render each page of the
website (the navbar is embedded within each individual HTML file). Thus you
could run the same command as above:

```{r wflow-publish-navbar}
wflow_publish("analysis/_site.yml", "Add main result page to navbar",
              republish = TRUE)
```

[navbar]: http://rmarkdown.rstudio.com/rmarkdown_websites.html#site_navigation

## Setup SSH keys

Using the https protocol to communicate with GitHub is tedious because it
requires entering your GitHub username and password. Using SSH keys for
authentication removes the password requirement. Follow these [GitHub
instructions][ssh] for creating SSH keys and linking them to your GitHub
account. You'll need to create separate SSH keys and link them each to GitHub
for each machine where you clone your Git repository.

After you create your SSH keys and add them to your GitHub account, you'll need
to instruct your local Git repository to use the SSH protocol. For a
hypothetical GitHub username of "myname" and GitHub repository of "myproject",
you would change the remote "origin" (the default name by convention) using the
function `wflow_git_remote()`:

```{r https-to-ssh}
wflow_git_remote(remote = "origin", user = "myname", repo = "myproject",
              protocol = "ssh", action = "set_url")
```

Alternatively you could update the remote URL using Git directly in the shell.
See this GitHub documentation on [changing a remote URL][set-url] for
instructions.

[ssh]: https://help.github.com/articles/generating-an-ssh-key/
[set-url]: https://help.github.com/articles/changing-a-remote-s-url/

## Change the session information function

The default function used to report the session information is `sessionInfo()`.
To change this, you can edit this setting in `_workflowr.yml`. For example, to
instead use `devtools::session_info()`, add the following line to
`_workflowr.yml`:

```
sessioninfo: "devtools::session_info()"
```

If you'd prefer to manually insert a more complex report of the session
information, disable the automatic reporting by adding the following to
`_workflowr.yml`:

```
sessioninfo: ""
```

Note however that workflowr will still check for the presence of a session
information function. Specifically it expects to find either `sessionInfo` or
`session_info` somewhere in the R Markdown document.
