---
title: "Hosting workflowr websites using GitLab"
subtitle: "workflowr version `r utils::packageVersion('workflowr')`"
author: "Luke Zappia, John Blischak"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Hosting workflowr websites using GitLab}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## What is in this vignette?

By default workflowr assumes that the project will be hosted on GitHub, but this
is not always the case. Users may prefer to use another service or have a
private Git repository hosting instance. This vignette details how to host a
workflowr project on GitLab. Unlike GitHub Pages, GitLab Pages offers both
public and private sites. For more details, see the documentation for [GitLab
Pages][gitlab-pages]. Similar steps will be required for other platforms but
some of the specifics will be different.

[gitlab-pages]: https://docs.gitlab.com/ee/ci/yaml/README.html#pages

## Step 0: Set up a project

The first thing we need to do is set up the project we want to host. We can do
this by following the first few steps of the instructions in the "Getting
started" vignette. When you get to the section [Deploy the
website](wflow-01-getting-started.html#deploy-the-website), follow the rest of
the steps in this vignette.

## Step 1: Create a remote repository on GitLab

**Note:** You can skip this step if you'd like because GitLab will automatically
create the new repository after you push it in Step 4 below. This
[feature][push-to-create-a-new-project] was introduced in [GitLab
10.5][gitlab-10.5], released in February 2018.

[push-to-create-a-new-project]: https://docs.gitlab.com/ee/gitlab-basics/create-project.html#push-to-create-a-new-project
[gitlab-10.5]: https://about.gitlab.com/2018/02/22/gitlab-10-5-released/

Log in to the GitLab instance you want to use and create a repository to host
your project (follow this [link](https://gitlab.com/projects/new) if you are
using GitLab.com). We recommend setting the project to be Public so that others
can inspect the code behind your results and extend your work.

## Step 2: Configure your local workflowr project to use GitLab

You will need to know your user name and the repository name for the following
steps (here we are going to use "myname" and "myproject") as well as a URL for
the hosting instance. The example below assumes you are using GitLab.com. If
instead you are using a custom instance of GitLab, you will need to change the
value for the argument `domain` accordingly ^[For example, the University of
Chicago hosts a GitLab instance for its researchers at
https://git.rcc.uchicago.edu/, which would require setting `domain =
"git.rcc.uchicago.edu"`].

```{r wflow-use-gitlab, eval=FALSE}
wflow_use_gitlab(username = "myname", repository = "myproject")
```

The function `wflow_use_gitlab()` automates all the local configuration
necessary to use GitLab. It changes the website directory from `docs/` to
`public/`, it creates the GitLab-specific configuration file `.gitlab-ci.yml`
with the necessary settings, and it connects the local Git repository to
communicate with the remote repository on GitLab.

## Step 3: Republish the analyses

In order for the correct URLs to past versions to be inserted into the HTML
pages, republish the analyses with `wflow_publish()`.

```
wflow_publish(republish = TRUE)
```

## Step 4: Push to GitLab

As a final step, push the workflowr project to GitLab (you will be prompted for
your GitLab username and password):

```{r wflow-git-push, eval=FALSE}
wflow_git_push()
```

If this step has worked correctly you should be able to refresh your GitLab page
and see all the files in your workflowr project. You can view your site at
`myname.gitlab.io/myproject/`, replacing with your username and project (note it
may take a minute for the site to be deployed).

If you skipped Step 1 above, the new repository created during the initial push
will be private by default. Unless you are working with sensitive data, you
should consider making the project public so that it is easier to share with
other researchers (e.g. collaborators, reviewers). You can change the visibility
by going to `Settings` -> `General` -> `Visibility` and changing `Project
visibility` to `Public`.

## Access control for private sites

If you need to keep your project private, you can [grant access][access-control]
to your collaborators by going to `Settings` -> `Members`. You can invite them
to the project via email, but they'll need a GitLab login to access the source
code and site. They can login to GitLab using common social sites like Google
and Twitter.

[access-control]: https://gitlab.com/help/user/project/pages/pages_access_control.md

## Compatibility with custom GitLab instances

Currently workflowr works best with the public GitLab instance hosted at
gitlab.com. If you are using a custom GitLab instance that is hosted by your
institution, it may not work as smoothly.

If you cannot view your workflowr website, this may be because your
administrators have not enabled [GitLab Pages][gitlab-pages]. You will need to
email them to activate this feature. You can include this link to the [GitLab
Pages administration][gitlab-pages-admin] instructions.

[gitlab-pages-admin]: https://git.rcc.uchicago.edu/help/administration/pages/index.md

If GitLab Pages is enabled, the links to past versions of the R Markdown files
should work correctly (open an [Issue][workflowr-issues] if you are having
problems). However, there is currently no way to conveniently view the past
versions of the HTML files. This is because workflowr uses the free service
[raw.githack.com][] to host past HTML files, and it only supports the URLs
`raw.githubusercontent.com`, `gist.githubusercontent.com`, `bitbucket.org`, and
`gitlab.com`.

[raw.githack.com]: http://raw.githack.com/
[workflowr-issues]: https://github.com/jdblischak/workflowr/issues
